Interactive Filter and Display of Hillary Clinton's Emails: \\A Cautionary Tale of Metadata

### 2017-04-10 VAST_submission/vgtc_journal_latex/paper.tex
Cleaning up text

Some sense of the content of the emails can be had from the updated term frequency displays shown in Figure~\ref{fig:TermFreqB1LibyaBuildup}.
\begin{figure}[h]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.4\linewidth]{TFIDFB1LibyaBuildup} &
$~~~~$ &
\includegraphics[width=0.4\linewidth]{TermFreqB1LibyaBuildup} 
\end{tabular}
\caption{Frequent terms: B1 only emails, Feb. 25 -- March 18, 2011}
\label{fig:TermFreqB1LibyaBuildup}
\end{center}
\end{figure}
Not surprisingly,  Libya and its leader figure prominently in this restricted set of emails.

Perhaps more surprising is a term like ``windrush''.  A little investigation shows that this is a reference to ``Windrush Ventures'' (e.g. see  \cite{windrushTelegraph, windrushGuardian}),  a company owned by former U.K. Prime Minister Tony Blair .  ``Windrush'' appears at the end of emails as part of Mr. Blair's electronic signature.   Mr. Blair is part of this B1 selected correspondence.

@@

\subsection{An email gap?}
\label{sect:emailGap}
Towards the very right of Figure~\ref{fig:lastTwoYears} there appears to be an anomaly in both time series.  There is a noticeable flat area at zero in both series.  If we move the leftmost time sliders to the June 1, 2012, and the rightmost to the end, February 1, 2013, we have the email volumes for the last 8 months of the archive.  The resulting series are shown in Figure~\ref{fig:EmailVolumeB1Jun1On}.
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.95\linewidth]{EmailVolumeB1Jun1On}
\caption{June 1, 2012 -- February 1, 2013:  Selected Emails are those that are B1 redacted}
\label{fig:EmailVolumeB1Jun1On}
\end{center}
\end{figure}
There appears to be no emails for a fairly large chunk of time.  In fact, there are few or no emails in the archive from October 30 to November 9, 2012.  This seems curious, given that the archive is supposed to be complete in Secretary Clinton's work-related email during this time period.  
The block of no email appears as an even more unusual pattern in Figure~\ref{fig:teaser}(a).  There the block is conspicuous by the absence amongst the emails from September 11 to November 23, 2012.  The email on which the displays of Figure~\ref{fig:teaser} are based, is no longer filtered by FOIA code.  It contains all email in the archive in this time period.  Direct examination of the archives at Wikileaks and the State Department show no mail in this empty block beyond that seen in Figure~\ref{fig:teaser}(a).

@@

One might ask what is happening around this time period?  In a word, Benghazi.  Five days after the assault, the U.S. Ambassador to the United Nations, Susan Rice, made a series of television appearances explaining that the attack on the U.S. Benghazi mission was a spontaneous response a  ``hateful video''  which had hours earlier caused violent protests in Cairo against the U.S. Embassy \cite{faceTheNationSept16}.   Four years later, it will be revealed as part of the House Select Committee on Benghazi report \cite{BenghaziReport} that Secretary Clinton had as early as September 12, 2012, communicated to Egyptian Prime Minister Kandil that the attack was known to be planned by Al Qaeda affiliates, it was not a protest, and that it had nothing to do with the film \cite{ClintonAdmitsAttackSept12}.  By the end of September the U.S. administration was calling it a terrorist attack \cite{factCheckerBenghazi}.  A month after the attack the administration was defending itself against charges that it deliberately downplayed a terrorist attack in Libya for political reasons in a presidential election year \cite{susanRiceWashPost2012}.   In October, focus shifted to the security at Benghazi and whether it had been reduced prior to the attack  \cite{cbsSecurity}.  The House Committee on Oversight \& Government Reform began looking into the security aspect \cite{cbsNewsHouseOversight2012} in earnest and found evidence that security issues were known and that requests from the U.S. Embassy in Libya for additional security personnel had been turned down by the State Department \cite{BenghaziSecurityTestimony}.  On October 26, CIA operatives who had defended the Benghazi mission against the terrorist attacks came forward with their story of what happened, on the ground.  They maintain that it was an organized attack and tell reporters that they were told to ``stand down'' \cite{standDownOrder}.  From October 26 to November 9 2012, no email appears in the archive as being sent from Secretary Clinton.

@@

To explore this possibility,  the time slider can be adjusted to a relatively small span, say of thirty days.  This would allow gaps in email of smaller sizes to be detected.  To focus attention on Secretary Clinton's behaviour, only those emails sent from Secretary Clinton would be displayed.  
Moving the slider, from left to right, from early 2009 to early 2013, allows one to move a one month wide magnifying glass across the whole of the archive.  A few contiguous days with no email from Clinton will appear as a small horizontal line at zero in the email volume (magenta coloured and lower) time series.  Once found, the sliders can be further narrowed to identify the dates of the patch.

@@

%% if specified like this the section will be committed in review mode
\acknowledgments{
This work was supported in part by
a Discovery grant from the Natural Sciences and Engineering Research Council of Canada..}

>>>>

Some sense of the content of the emails can be had from the updated term frequency displays shown in Figure~\ref{fig:TermFreqB1LibyaBuildup}.
\begin{figure}[h]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.4\linewidth]{TFIDFB1LibyaBuildup} &
$~~~~$ &
\includegraphics[width=0.4\linewidth]{TermFreqB1LibyaBuildup} 
\end{tabular}
\caption{Frequent terms: B1 only emails, Feb. 25 -- March 18, 2011}
\label{fig:TermFreqB1LibyaBuildup}
\end{center}
\end{figure}
Not surprisingly,  Libya and its leader figure prominently in this restricted set of emails.  
Perhaps more surprising is a term like ``windrush''.  A little investigation shows that this is a reference to ``Windrush Ventures'' (e.g. see  \cite{windrushTelegraph, windrushGuardian}),  a company owned by former U.K. Prime Minister Tony Blair .  ``Windrush'' appears at the end of emails as part of Mr. Blair's electronic signature.   
Mr. Blair is part of this B1 selected correspondence.

@@

\subsection{An email gap?}
\label{sect:emailGap}
Towards the very right of Figure~\ref{fig:lastTwoYears} there appears to be an anomaly in both time series.  There is a noticeable flat area at zero in both series.

If we move the leftmost time sliders to the June 1, 2012, and the rightmost to the end, February 1, 2013, we have the email volumes for the last 8 months of the archive.  The resulting series are shown in Figure~\ref{fig:EmailVolumeB1Jun1On}.
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.95\linewidth]{EmailVolumeB1Jun1On}
\caption{June 1, 2012 -- February 1, 2013:  Selected Emails are those that are B1 redacted}
\label{fig:EmailVolumeB1Jun1On}
\end{center}
\end{figure}
There appears to be no emails for a fairly large chunk of time.  In fact, there are few or no emails in the archive from October 30 to November 9, 2012.  
This seems curious, given that the archive is supposed to be complete in Secretary Clinton's work-related email during this time period.

The block of no email appears as an even more unusual pattern in Figure~\ref{fig:teaser}(a).  There the block is conspicuous by the absence amongst the emails from September 11 to November 23, 2012.  The email on which the displays of Figure~\ref{fig:teaser} are based, is no longer filtered by FOIA code.  It contains all email in the archive in this time period.  Direct examination of the archives at Wikileaks and the State Department show no mail in this empty block beyond that seen in Figure~\ref{fig:teaser}(a).

@@

One might ask what is happening around this time period?  In a word, Benghazi.

Five days after the assault, the U.S. Ambassador to the United Nations, Susan Rice, made a series of television appearances explaining that the attack on the U.S. Benghazi mission was a spontaneous response a  ``hateful video''  which had hours earlier caused violent protests in Cairo against the U.S. Embassy \cite{faceTheNationSept16}.   Four years later, it will be revealed as part of the House Select Committee on Benghazi report \cite{BenghaziReport} that Secretary Clinton had as early as September 12, 2012, communicated to Egyptian Prime Minister Kandil that the attack was known to be planned by Al Qaeda affiliates, it was not a protest, and that it had nothing to do with the film \cite{ClintonAdmitsAttackSept12}.  By the end of September the U.S. administration was calling it a terrorist attack \cite{factCheckerBenghazi}.  A month after the attack the administration was defending itself against charges that it deliberately downplayed a terrorist attack in Libya for political reasons in a presidential election year \cite{susanRiceWashPost2012}.

In October, focus shifted to the security at Benghazi and whether it had been reduced prior to the attack  \cite{cbsSecurity}.  The House Committee on Oversight \& Government Reform began looking into the security aspect \cite{cbsNewsHouseOversight2012} in earnest and found evidence that security issues were known and that requests from the U.S. Embassy in Libya for additional security personnel had been turned down by the State Department \cite{BenghaziSecurityTestimony}.

On October 26, CIA operatives who had defended the Benghazi mission against the terrorist attacks came forward with their story of what happened, on the ground.  They maintain that it was an organized attack and tell reporters that they were told to ``stand down'' \cite{standDownOrder}.  From October 26 to November 9 2012, no email appears in the archive as being sent from Secretary Clinton.

@@

To explore this possibility,  the time slider can be adjusted to a relatively small span, say of thirty days length.  This would allow gaps in email of smaller sizes to be detected.  To focus attention on Secretary Clinton's behaviour, only those emails sent from Secretary Clinton will be displayed.  
Moving the slider, from left to right, from early 2009 to early 2013, allows one to move a one month wide magnifying glass across the whole of the archive.  A few contiguous days with no email from Clinton will appear as a small horizontal line at zero in the email volume (magenta coloured and lower) time series.  Once found, the sliders can be further narrowed to identify the dates of the patch.

@@

%% if specified like this the section will be committed in review mode
\acknowledgments{
This work was supported in part by
a Discovery grant from the Natural Sciences and Engineering Research Council of Canada.}
